---
title: "homework-2"
author: "Amina Mohammed (17301920), Joost Dijkstra (),Edward Tandia (17310806)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
```{r, echo = FALSE, include = FALSE, message = FALSE}
source(here::here("Setup.R"))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Problem 1. statement

Your boss would like that you analyze some business data. Let $X_1,\dots,X_n$ represents $n$ data points. Suppose you are interested in a statistic, say $\hat{\theta}$. In particular, we consider the mean, the median and the max. You are asked not only to report the statistics, but also an estimator of their variances.

In order to accomplish this task, you will implement the bootstrap. The bootstrap is a well known method in statistics since Efron's seminal paper in 1979. The bootstrap is easy to implement and straightforward to use. There exist many different schemes for the bootstrap, we present the simplest form:

1.  Compute the statistic on the sample: $\hat{\theta} = g(x_1,\dots,x_n)$.
2.  Create a new sample $x_1^\ast,\dots,x_n^\ast$ by drawing data from the original sample **at random with replacement**. This new sample is called a *bootstrapped sample*.
3.  Compute the statistic on the bootstrapped sample: $\hat{\theta}^\ast = g(x_1^\ast,\dots,x_n^\ast)$.
4.  Repeat 2. and 3. $B$ times.
5.  Compute the unbiased estimator of the variance: $$\frac{1}{B-1}\sum_{b=1}(\hat{\theta}^\ast_{b}-\hat{\theta})^2.$$

___Answers Problem 1___
```{r echo = FALSE, include = FALSE, message = FALSE}
supermarket <- read.csv("supermarket_sales.csv")
unitPrice <- as.matrix(supermarket[,7])

# statistic on the sample 
têta_hat_mean <- mean(unitPrice)       # mean of unit price (theoretical)
têta_hat_median <- median(unitPrice)   # median of unit price (theoretical)
têta_hat_max <- max(unitPrice)         # max of unit price (theoretical)

B <- 1000                     #number of iteration

```
In order to perfom the unbiased estimator of the variance, we're going to need the boostrapped statistic and the data statistic. Let's suppose $\hat{\theta}$ which can be either the mean, median and the maximum of the dataset.

Since Bootstrapping is the process of resampling with replacement, we're going to define the boostraped sample as follow: 


```{r echo = FALSE, include = FALSE, message = FALSE}

tableStatisticalvalue <- data.frame(Statistics= c("mean", "median", "max"), Value=c(têta_hat_mean,têta_hat_median,têta_hat_max))

```
```{r echo=FALSE}
kable(head(tableStatisticalvalue), booktabs = TRUE) %>%
  kable_styling(font_size = 16)
```


**Question 1.**  Load the dataset `supermarket_sales.csv`. We are interested in the column `Unit.price` and $\hat{\theta}$ is the mean. Using a `for` loop, compute the unbiased estimator of the variance with $B=1'000$.


```{r Bootstrap with for loop mean}
set.seed(123)
têta_hat_loop <- rep(0,B)        #save the unbiased estimator of var

# Bootstrap with a for loop 
for (i in 1:B) {
  
  têta_hat_loop[i] <- mean(sample(unitPrice, replace = TRUE))                      # bootstrap estimate of mean
}

estimator_loop <-(1/(B-1))*(sum(têta_hat_loop- têta_hat_mean)^2)
estimator_loop
```
**Answer 1 :** The unbiased estimator of the variance with $B=1'000$ is *estimator_loop*

**Question 2. ** Propose an alternative solution without any control structure (i.e., no `for` loop, `while` loop, ...). You can for example construct a matrix and use `colMeans()` or `rowMeans()` functions.

```{r Bootstrapping without control structure}
"library(bootstrap)
matrice <- matrix(ncol = B,nrow=B)

#matrice <-mcopy paste the colum btimes

set.seed(123)

#create a vector of teta_hat
theta <- function(d){colMeans(d)} 

#create a matrix with B time Unitprice
boots_noloop <- bootstrap(unitPrice,B, theta)

teta_hat_noloop <- boots_noloop$thetastar


estimator_noloop <-(1/(B-1))*(sum(teta_hat_noloop- têta_hat)^2)
estimator_noloop
"
```

**Answer 2 :** The unbiased estimator of the variance with $B=1'000$ is *estimator_noloop*

**Question 3**.  Compare the performances of your solutions at 1. and 2. by:

    **a.** Profiling your two implementations. Use `profvis::profvis({...})`. Comment on the comparison, in particular the memory usage and computation time.  

```{r }
#Profiling Bootstrap with for loop 

# Profiling Bootstrap with a for loop 
set.seed(123)

profvis::profvis({
têta_hat_loop <- rep(0,B)        #save the unbiased estimator of var

# Bootstrap with a for loop 
for (i in 1:B) {
  
  têta_hat_loop[i] <- mean(sample(unitPrice, replace = TRUE))                      # bootstrap estimate of mean
}

estimator_loop <-(1/(B-1))*(sum(têta_hat_loop- têta_hat_mean)^2)
estimator_loop
 })

```

```{r }
"
# Profiling Bootstrap without loop 
set.seed(123)
profvis::profvis({

  })
"
```

**Answer 3.a**:  The memory usage and computation time for the Bootstrap with control structure are : 
* memory usage :
* computation time:

And the memory usage and computation time for Bootstrap without control structure are: 
* memory usage :
* computation time : 

We take less time in the case of bootstrapping with / without control structure.

<br>

**b.** Benchmarking your two implementations. Use `microbenchmark::microbenchmark({...},{...}). Comment the results.

```{r}
"
microbenchmark::microbenchmark({
  têta_hat_loop <- rep(0,B)     
  for (i in 1:B) {
  
  têta_hat_loop[i] <- mean(sample(unitPrice, replace = TRUE))               
}

estimator_loop <-(1/(B-1))*(sum(têta_hat_loop- têta_hat_mean)^2)
estimator_loop},

{...})
"

```
**Answer 3.b**:

**Question 4.**  Repeat 2. with the *median* as the statistic of interest. You are still required not use any control structures. There are no `colMedians()` or `rowMedians()` functions. Instead, use the `apply` function. Comment the results in comparison with the mean.

```{r Bootstrap with apply function }
set.seed(123)

myBootFun <- function(d) {
    boot(d, median, R = 1000)
}

têta_hat_apply <- sapply(unitPrice, function(x) myBootFun(x)$t0 )

estimator_apply <-(1/(B-1))*(sum(têta_hat_apply- têta_hat_median)^2)
estimator_apply

```

Before anything else, let's remember that The median is sometimes used as opposed to the mean when there are outliers in the sequence that might skew the average of the values.In a normal distribution, the median is the same as the mean and the mode, which is not the case for us. 

Compare to the mean, the median value is `XXX` which is higher/lower that the mean which is at `XXX`. Indeed, A lower median mean, 



**Question 5.**  Repeat 1. with the *max* as the statistic of interest. What do you observe? Comment.
```{r Bootstrap with for loop max }

set.seed(123)
têta_hat_loop_max <- rep(0,B)        #save the unbiased estimator of var

# Bootstrap with a for loop 
for (i in 1:B) {
  
  têta_hat_loop_max[i] <- max(sample(unitPrice, replace = TRUE))                      # bootstrap estimate of mean
}

estimator_loop_max <-(1/(B-1))*(sum(têta_hat_loop_max- têta_hat_max)^2)
estimator_loop_max
```

**Answer 5:** In this case the statistic of interest is the maximum, as a reminder we had the following observation:
* unbiased estimator of var with mean boostraping : 0.6247326
* unbiased estimator of var with max boostraping : 0.07403403

We can observe that we have a decrease on the variance. 

```{r echo = FALSE, include = FALSE, message = FALSE}
df_answer5 <- data.frame (Statistic  = c("mean", "max"),
                  "têta" = c(têta_hat_mean, têta_hat_max),
                  "Unbiased Estimator" = c(estimator_loop,estimator_loop_max)
                  )
```
```{r echo=FALSE}
df_answer5 %>%
  kable(booktabs = TRUE) %>%
  kable_styling(font_size =16)
```